---
title: '`r toupper(gsub("_.*$", "", basename(getwd())))`'
author: Frank Zhang
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
    html_document:
        css: css/style.css
---

```{r, eval=FALSE, echo=FALSE}
rmarkdown::render(paste(gsub("_.*$", "", basename(getwd())), ".Rmd", sep = ""), output_dir = "dist")
```

```{r}
reticulate::use_virtualenv(getOption("venv"))
```

```{r}
# don't run the line below, it fucks up later imports for some reason
# py <- import("__builtin__")
modules::import(gsubfn)
modules::import(foreach)
modules::import(magrittr)
py <- reticulate::import("builtins")
# tools <- reticulate::import("rpytools")
np <- reticulate::import("numpy")
mx <- reticulate::import("mxnet")
```

This is an example of how to define a python class in reticulate. It will come in handy later.

For some reason, this is not always reliable.

- It seems like the first invocation of type always returns a `class '*'` instance, that fails when invoked. However, after failing, all instances of py$type return a `class 'rpytools.call.*'`, which does work.
- Ended up giving up finding a way to convert the class and wrote this preamble to fail first quietly.

```{r}
tryCatch({
    py$type("_", reticulate::tuple(), reticulate::dict(
      "__init__" = function(self, x) {
        reticulate::py_set_attr(self, "_", x)
        return(NULL)
      }
  ))(NULL)
}, error = function (e) {i<-1})
```

```{r}
Test <- py$type("Test", reticulate::tuple(), reticulate::dict(
  "__init__" = function(self, x) {
    reticulate::py_set_attr(self, "a", x)
    return(NULL)
  }
))
test <- Test(1)
test$a
```

```{r}
Test1 <- py$type("Test1", reticulate::tuple(Test), reticulate::dict(
  "__init__" = function(self, x) {
    Test$`__init__`(self, x)
    return(NULL)
  }
))
test1 <- Test1(5)
test1$a
```

We’ll also want to set the contexts for our data and our models.

```{r}
data_ctx <- mx$cpu()
model_ctx <- mx$cpu()
```

Some metadata.

```{r}
epochs <- py$int(10)
batch_size <- py$int(64)
num_inputs <- py$int(784)
num_outputs <- py$int(10)
num_examples <- py$int(60000)
```

A transform function.

- Notes:
    - For functions to return multiple arguments, have to use reticulate::tuple, list won't work
    - Arithmetic operators won't work on mxnet data that hasn't been converted (things like +,-,*,/)
        - Instead, use the `__add__`, `__div__` methods
        - For mxnet data of a certain dimension (at least one dimension == 1), the astype method won't work (or any method for that matter as it will complain about atomic vectors)
            - Luckily, normal operators will work on these
            - And reticulate `np_array` actually works on them, though not more complicated data

```{r}
train_data <- mx$gluon$data$DataLoader(
    mx$gluon$data$vision$MNIST(
        train = TRUE,
        transform = function(data, label) {
            reticulate::tuple(
                data$astype(np$float32)$`__div__`(255),
                reticulate::np_array(label, "float32")
            )
        }
    ),
    batch_size, shuffle = TRUE
)

test_data <- mx$gluon$data$DataLoader(
    mx$gluon$data.vision$MNIST(
        train = FALSE,
        transform = function(data, label) {
            reticulate::tuple(
                data$astype(np$float32)$`__div__`(255),
                reticulate::np_array(label, "float32")
            )
        }
    ),
    batch_size, shuffle = FALSE
)
```

```{r}
net <- mx$gluon$nn$Sequential()
with(
    net$name_scope(),
    {
        net$add(mx$gluon$nn$Dense(py$int(64), activation = "relu"))
        net$add(mx$gluon$nn$Dense(py$int(64), activation = "relu"))
        net$add(mx$gluon$nn$Dense(num_outputs))
    }
)
```

Let's try to get the data to show that everything works.

```{r}
reticulate::iter_next(py$iter(train_data))
```

Parameter initialization

```{r}
net$collect_params()$initialize(mx$init$Normal(sigma = 0.1), ctx = model_ctx)
```

Softmax cross entropy loss

```{r}
softmax_cross_entropy <- mx$gluon$loss$SoftmaxCrossEntropyLoss()
```

Accuracy

```{r}
acc <- mx$metric$Accuracy()
```

Optimizer

```{r}
trainer <- mx$gluon$Trainer(net$collect_params(), "sgd", reticulate::dict(
    "learning_rate" = 0.01
))
```

Simulate a batch through the network

- Different methods of method chaining with magrittr are shown, the problem is that the data object is always passed as the first argument and is unexpected in python method chains
    - The first method doesn't raise linter errors but has more characters
    - The second method is shorter but linter complains about the {}'s

```{r}
list[data, label] <- reticulate::iter_next(py$iter(train_data))

data <- data %>% (function(.)
    .$as_in_context(model_ctx)) %>% (function(.)
    .$reshape(c(py$int(-1), py$int(784))))

label <- label %>%
    {.$as_in_context(model_ctx)}

output <- net(data)
# expand dim is necessary because it comes out as n, as opposed to n,1 dimension
predictions <- mx$nd$argmax(output, axis = py$int(1))$expand_dims(axis=py$int(1))
acc$update(preds = predictions, labels = label)
acc$get()
```

Simulate a training batch through the network

```{r}
list[data, label] <- reticulate::iter_next(py$iter(train_data))

data <- data %>% (function(.)
    .$as_in_context(model_ctx)) %>% (function(.)
    .$reshape(c(py$int(-1), py$int(784))))

label <- label %>% (function(.)
    .$as_in_context(model_ctx))

with(mx$autograd$record(), {
    output <- net(data)
    loss <- softmax_cross_entropy(output, label)
})

loss$backward()
trainer$step(data$shape[[1]])
mx$nd$sum(loss)$asscalar()
```

Messing around with dataset and dataloader classes to get a feel for the api.

```{r}
dataset <- mx$gluon$data$dataset$ArrayDataset(
    mx$random$uniform(shape = list(py$int(10), py$int(3))),
    mx$random$uniform(shape = list(py$int(10), py$int(1)))
)
```

These two implementations are equivalent, obviously the second one is better since we don't have to invoke a hidden method.

- similar thing applies to `__iter__`, luckily, `py$iter` returns the iterator.

```{r}
dataset$`__getitem__`(py$int(1))
```
```{r}
reticulate::py_get_item(dataset, py$int(1))
```
```{r}
data_loader <- mx$gluon$data$DataLoader(dataset, batch_size = 5)
```

Calling `py$iter` on the data loader object will regenerate the data loader for another iteration. Otherwise, it will be empty after it runs once through.

```{r}
test <- py$iter(data_loader)
```

This is how to get tuple unpacking and enumerate functionality from the reticulate iterator

- reticulate will return NULL by default if iterator = empty, iterators doesn't stop for nulls so need a check to call `stop("StopIteration")` to halt the iterators iter
- `icount` gets the enumerate like index
- tuple unpacking accomplished by `list[]` is provided by the `gsubfn` library
    - must be called outside `foreach` in the `do` body, doesn't work inside
- self invoking function is provided to provide variable proper scoping

```{r}
(function() {
foreach(
    D = iterators::iter(function() reticulate::iter_next(test) %T>%
        (. %>% is.null %>% if (.) stop("StopIteration", call. = FALSE))
    ),
    i = iterators::icount()
) %do% {
    list[X_batch, y_batch] <- D
    print(i)
    y_batch$shape
}
})()
```





