---
title: '`r toupper(gsub("_.*$", "", basename(getwd())))`'
author: Frank Zhang
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
    html_document:
        css: css/style.css
---

```{r, eval=FALSE, echo=FALSE}
rmarkdown::render(paste(gsub("_.*$", "", basename(getwd())), ".Rmd", sep = ""), output_dir = "dist")
```

Turns out as.integer also works for reticulate, by default, R doesn't use ints.

```{r}
reticulate::use_virtualenv(getOption("venv"))
mx <- reticulate::import("mxnet")
py <- reticulate::import("builtins")
modules::import(magrittr)
mx$random$seed(1 %>% as.integer)
```

Empty grabs memory and hands back matrix without initialization. Also turns out c() will work just as well as list() for some things.

```{r}
mx$nd$empty(c(3 %>% as.integer, 4 %>% as.integer))
```

Matrix of zeros.

```{r}
mx$nd$zeros(c(3 %>% as.integer, 4 %>% as.integer))
```

Matrix of ones.

```{r}
mx$nd$ones(c(3 %>% as.integer, 4 %>% as.integer))
```

Matrix of standard normal.

```{r}
mx$nd$random_normal(0, 1, shape = c(3 %>% as.integer, 4 %>% as.integer))
```

Getting the shape. Turns out it returns a list and not a vector.

```{r}
mx$nd$empty(c(3 %>% as.integer, 4 %>% as.integer))$shape
```

Query size, which is the number of elements.

```{r}
mx$nd$empty(c(3 %>% as.integer, 4 %>% as.integer))$size
```

Addition, the + operator from R will not work.

```{r}
mx$nd$zeros(c(3 %>% as.integer, 4 %>% as.integer))$`__add__`(mx$nd$ones(c(3 %>% as.integer, 4 %>% as.integer)))
```

Likewise with * multiplication.

```{r}
mx$nd$zeros(c(3 %>% as.integer, 4 %>% as.integer))$`__mul__`(mx$nd$ones(c(3 %>% as.integer, 4 %>% as.integer)))
```

Exponentiation.

```{r}
mx$nd$exp(mx$nd$ones(c(3 %>% as.integer, 4 %>% as.integer)))
```

Transpose and dot product.

```{r}
mx$nd$dot(mx$nd$ones(c(3 %>% as.integer, 4 %>% as.integer)), mx$nd$ones(c(3 %>% as.integer, 4 %>% as.integer))$T)
```

In place operations. The first block shows how a new array is produced and overwrites the first.

```{r}
x <- mx$nd$random_normal(0, 1, shape = c(3 %>% as.integer, 4 %>% as.integer))
y <- mx$nd$zeros(c(3 %>% as.integer, 4 %>% as.integer))
print(paste("id(y)", py$id(y)))
y <- y$`__add__`(x)
print(paste("id(y)", py$id(y)))
```

Assign result to previously allocated array with slice notation. Turns out slicing with mxnet in reticulate defaults to 0-indexing.

```{r}
x
```

This kind of subsetting will not support later assignment. Also doesn't even work with columns. (Unless you provide it `reticulate::tuple`).

```{r}
x[(1:3)-1]
```

Apparently, neither will this subsetting allow assignment.

```{r}
x$slice(c(0 %>% as.integer, NULL), c(x$shape[[1]] %>% as.integer, NULL))
```

Couldn't manage to get slice assignment to work, but copyto exists.

```{r}
x <- mx$nd$random_normal(0, 1, shape = c(3 %>% as.integer, 4 %>% as.integer))
y <- mx$nd$zeros(c(3 %>% as.integer, 4 %>% as.integer))
print(y)
print(paste("id(y)", py$id(y)))

y %>% (function(.)
    .$`__add__`(x)) %>% (function(.)
    .$copyto(y)
)

print(y)
print(paste("id(y)", py$id(y)))
```

The above example is obvious that x+y will still allocate a temporary buffer to store result before copying it over to y. Can use elemwise_add to avoid temporary buffers.

```{r}
mx$nd$elemwise_add(x, y, out = y)
```

There is also in-place arithmetic.

```{r}
print(x)
print(paste("id(x)", py$id(x)))
x$`__iadd__`(y)
print(x)
print(paste("id(x)", py$id(x)))
```

More slicing.

```{r}
x
```

```{r}
x[0:1]
```

Assignment doesn't work with subsetting, use `__setitem__` instead.

```{r}
x[0]$`__setitem__`(1 %>% as.integer, 5)
x
```

The indexing that R lets you do is just on the first dimension, it won't let you use commas to access other dimensions (complains about unused argument).

```{r}
x[1] %>% (function(.)
    .$`__setitem__`(1:2 %>% as.integer, 5))
x
```

Probably the best way to get a specific column.

```{r}
x$slice_axis(axis = 1 %>% as.integer, begin = 1 %>% as.integer, end = 2 %>% as.integer)
```

For some reason, setitem with slice axis just doesn't work. Slice is probably making a copy and setitem is doing the modification in place, so the original item doesn't get modified.

```{r}
print(x$slice_axis(axis = 1 %>% as.integer, begin = 1 %>% as.integer, end = 2 %>% as.integer) %T>% (function(.) .$`__setitem__`(2 %>% as.integer, 5)) %>% (function(.) print(.)))
print(x)
```

For some reason, this doesn't work either.

```{r}
x %>%
 .[reticulate::tuple(1:.$shape[[1]] - 1 %>% as.integer, 1 %>% as.integer)] %T>% (function(.)
 .$`__setitem__`(1:.$size - 1 %>% as.integer, 5)) %>% (function(.)
 .$`__getitem__`(1:.$size - 1 %>% as.integer))
print(x)
```

This also doesn't work.

```{r}
x %>%
 .[reticulate::tuple(1:.$shape[[1]] - 1 %>% as.integer, 1 %>% as.integer)] %T>% (function(.)
 .$`__setitem__`(1:.$size - 1 %>% as.integer, 5)) %>% (function(.)
 .$`__getitem__`(1:.$size - 1 %>% as.integer)) %>% (function(.)
 .$copyto((
     x %>%
      .[reticulate::tuple(1:.$shape[[1]] - 1 %>% as.integer, 1 %>% as.integer)]
 )))
print(x)
```

This finally works, have to use reticulate `py_set_item` to interface correctly with object reference. Also need to use reticulate tuple to allow multidimensional indexing since it only takes one argument. Note that c and list do not work.

```{r}
x %>%
    reticulate::py_set_item(
        reticulate::tuple(
            1:.$shape[[1]] - 1 %>% as.integer,
            1 %>% as.integer
        ),
        2
    )
print(x)
```

So in summary, `x[number]` will work with setitem for assignment, but that notation only works with rows (i.e. first dimension). When trying to use that notation with a tuple to access a column, setitem will no longer work because it won't modify the original object.

Some examples on broadcasting follow. Also, assignment with () will no longer be silent.

```{r}
(x <- mx$nd$ones(c(3 %>% as.integer, 3 %>% as.integer)))
(y <- mx$nd$arange(3 %>% as.integer))
(x$`__add__`(y))
```

```{r}
(y <- y$reshape(c(3 %>% as.integer, 1 %>% as.integer)))
(x$`__add__`(y))
```

Converting to numpy array and creation of a new nd array from numpy array.

```{r}
x$asnumpy() %T>% {py$type(.) %>% print()} %>% mx$nd$array()
```

You can also control the context of where arrays are created.

```{r}
mx$nd$ones(c(3 %>% as.integer, 3 %>% as.integer), ctx = mx$cpu(1 %>% as.integer))
```

To move arrays, you can use `copyto`.

```{r}
mx$nd$ones(c(3 %>% as.integer, 3 %>% as.integer), ctx = mx$cpu(1 %>% as.integer))$copyto(mx$cpu(0 %>% as.integer))
```

In order to perform an operation, all the arrays have to be in the same context. Note that copying to a different context is a slow operation. You can check the context.

```{r}
mx$nd$ones(c(3 %>% as.integer, 3 %>% as.integer), ctx = mx$cpu(1 %>% as.integer))$copyto(mx$cpu(0 %>% as.integer))$context
```

Copying to the same context that an array already exists will create a copy and reassign the array (this changes the id). Calling `as_in_context` will prevent this copying process if the array already exists.

```{r}
(z <- mx$nd$ones(c(3 %>% as.integer, 3 %>% as.integer), ctx = mx$cpu(1 %>% as.integer)))
print(py$id(z))
z <- z$copyto(mx$cpu(1 %>% as.integer))
print(py$id(z))
z <- z$as_in_context(mx$cpu(1 %>% as.integer))
print(py$id(z))
```