---
title: '`r toupper(gsub("_.*$", "", basename(getwd())))`'
author: Frank Zhang
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
    html_document:
        css: css/style.css
---

```{r, eval=FALSE, echo=FALSE}
rmarkdown::render(paste(gsub("_.*$", "", basename(getwd())), ".Rmd", sep = ""), output_dir = "dist")
```

Turns out as.integer also works for reticulate, by default, R doesn't use ints.

```{r}
reticulate::use_virtualenv(getOption("venv"))
mx <- reticulate::import("mxnet")
nd <- reticulate::import("mxnet")$nd
py <- reticulate::import("builtins")
modules::import(magrittr)
mx$random$seed(1 %>% as.integer)
```

Some basic scalar arithmetic.

```{r}
(x <- nd$array(list(3)))
(y <- nd$array(list(2)))
x$`__add__`(y)
x$`__mul__`(y)
x$`__div__`(y)
nd$power(x, y)
x$asscalar()
```

Vector creation and indexing, length and shape

```{r}
(u <- nd$arange(4))
u[3]
py$len(u)
u$shape
```

Some matrix examples, note that indexing only takes one argument, and if a number, will go in each index in order, hence it is impossible to select all columns for instance with this notation.

```{r}
(A <- nd$zeros(c(5 %>% as.integer, 4 %>% as.integer)))
(x <- nd$arange(20))
(A <- x$reshape(c(5 %>% as.integer, 4 %>% as.integer)))
A$T
A[2][3]
```

This is how to do it if you want to grab row 2 vs column 3. Remember that assignment with mx native `setitem` will not work with the second one.

```{r}
A[2]
A %>%
    .[reticulate::tuple(
        1:.$shape[[1]] - 1 %>% as.integer,
        3 %>% as.integer
    )]
```

A tensor.

```{r}
(A <- nd$arange(24)$reshape(c(2 %>% as.integer, 3 %>% as.integer, 4 %>% as.integer)))
```

The sum function, which collapses all elements into a single number.

```{r}
u
nd$sum(u)
A
nd$sum(A)
```

The mean function.

```{r}
A
nd$mean(A)
nd$sum(A)$`__div__`(A$size)
A %>% (function(.)
    nd$sum(.)$`__div__`(.$size))
```

Dot product.

```{r}
u
(v <- nd$ones_like(u)$`__mul__`(2))
nd$dot(u, v)
```

L2 norm and L1 norm.

```{r}
nd$norm(u)
nd$sum(nd$abs(u))
u %>% nd$abs() %>% nd$sum()
```