---
title: '`r toupper(gsub("_.*$", "", basename(getwd())))`'
author: Frank Zhang
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
    html_document:
        css: css/style.css
---

```{r, eval=FALSE, echo=FALSE}
rmarkdown::render(paste(gsub("_.*$", "", basename(getwd())), ".Rmd", sep = ""), output_dir = "dist")
```

In this example, the code that was used employed python2 print statements, so the default python was used instead of venv. Also, all reticulate modules were called with convert = FALSE to prevent having to think about numpy specifics.

```{r}
# reticulate::use_virtualenv(getOption("venv"))
np <- reticulate::import("numpy", convert = FALSE)
modules::import(magrittr)
modules::import(gsubfn)
```

Examples of how to use `import_from_path`. The filename is the module, and the path is the directory where the file is located. Specific elements can be imported with $.

```{r}
load_data <- reticulate::import_from_path("mondrianforest_utils", path = "./mondrianforest/src", convert = FALSE)$load_data

process_command_line <- reticulate::import_from_path("mondrianforest", path = "./mondrianforest/src", convert = FALSE)$process_command_line

MondrianForest <- reticulate::import_from_path("mondrianforest", path = "./mondrianforest/src", convert = FALSE)$MondrianForest

precompute_minimal <- reticulate::import_from_path("mondrianforest_utils", path = "./mondrianforest/src", convert = FALSE)$precompute_minimal

get_data_min_max <- reticulate::import_from_path("mondrianforest", path = "./mondrianforest/src", convert = FALSE)$get_data_min_max
```

Try out the mondrian on the dummy dataset, which is fast. The data path had to be adjusted to account for code invocation from this notebook.

```{r}
settings <- process_command_line()
reticulate::py_set_attr(settings, "data_path", "./mondrianforest/process_data/")
# reticulate::py_set_attr(settings, "dataset", "satimage")
# reticulate::py_set_attr(settings, "n_minibatches", 10 %>% as.integer)
reticulate::py_set_attr(settings, "debug", TRUE)
```

Load the dummy data.

```{r}
data <- load_data(settings)
```

```{r}
data
```

Instantiate instance of mondrian forest.

```{r}
mf <- MondrianForest(settings, data)
```

An example of how to do tuple instantiation with reticulate. If not automatic converting, need to explicitly call `py_to_r`. Also need `gsubfn`.

```{r}
list[param, cache] <- precompute_minimal(data, settings) %>% reticulate::py_to_r()
```

Call to fit the mondrian forest.

```{r}
mf$fit(data, data$train_ids_partition$current %>% reticulate::py_get_item(0), settings, param, cache)
```

How to access a dictionary with a number. The `0` notation won't work with python objects.

```{r}
data$train_ids_partition$current %>% reticulate::py_get_item(0)
```

Analyze the real sample dataset.

```{r}
settings <- process_command_line()
reticulate::py_set_attr(settings, "data_path", "./mondrianforest/process_data/")
reticulate::py_set_attr(settings, "dataset", "satimage")
reticulate::py_set_attr(settings, "n_minibatches", 10 %>% as.integer)
reticulate::py_set_attr(settings, "debug", TRUE)
```

```{r}
data <- load_data(settings)
```

Determine what `get_data_min_max` actually does. It seems to look at data and pull out each dimension max and min over the entire data.

```{r}
get_data_min_max(data, data$train_ids_partition$current %>% reticulate::py_get_item(0))[1]
```

